Estimating duration using survival analysis

### 2023-12-15 cis-perfect-testing.tex
Most simulation results for imperfect testing
First version of full text on imperfect testing
Full first draft of imperfect chapter

\subsection{Approach}

@@

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$.
  Otherwise, define $l_i^{(b)}$ and $r_i^{(b)}$ as the first and last days $b_i$ could have occurred.
  $l_i^{(b)}$ is the day after the last negative test prior to the first positive test in the episode.
  $r_i^{(b)}$ is the day of the first positive test.
  Similarly, define $l_i^{(e)}$ and $r_i^{(e)}$ as the first and last days $e_i$ could have occurred.
  If $i \in \set{D}$ then $n_{ia} =1$ admissible episodes occurred in individuals with test schedule $\sched{i}$, otherwise $n_{ia} = 0$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \theta)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i^C$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these observed in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is undetected with probability $p_{iu} = \prob((b, e) \in \Omega^C_i \mid \theta)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  This is all remaining episodes not in the previous sets.
  $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, it is inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \theta)$.
\end{itemize}

@@

Next, I derive $p_{ia}$.
By definition (see~\cpageref{perf-test:def:admissible}), we have:
\begin{align}
p_{ia} &= \prob((b, e) \in \alpha_i \mid \theta) \\
\alpha_i &= \{ (b, e) : l_i^{(b)} \leq b \leq r_i^{(b)} \wedge l_i^{(e)} \leq e \leq r_i^{(e)}\}.
\intertext{Suppressing the conditioning on $\theta$, this gives:}
p_{ia}
&= \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)}, l_i^{(e)} \leq E_i \leq r_i^{(e)} \right) \\
&= \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \\
&=\sum_{b = l_i^{(l)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid B_i = b \right) \prob \left(B_i = b \right) \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} - b + 1 \leq D_i \leq r_i^{(e)} - b + 1 \right) \prob \left(B_i = b \right) &\text{by def of $D_i$} \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right) \prob \left(B_i = b \right) &\text{by def of $S_\theta$} \\
&\propto \sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.

Next, I derive $p_{iu} = \prob((B_i, E_i) \in \Omega^C_i \mid \theta)$, the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega^C_i$ can be written as:
\begin{align}
\Omega_i^C = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Only $1 - p_u = 1 - \frac{1}{N} \sum_i^{\Ncis} p_{iu} = \frac{1}{N} \sum_i^{\Ncis} (1 - p_{iu})$ is required (see \cref{perf-test:eq:full-posterior-simplified}).
Therefore, I derive $1 - p_{iu}$, again suppressing the conditioning on $\theta$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_i, E_i) \in \Omega^C_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) + B_i > E_i \vee B_i \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) > D_i , B_i > \min(\sched_i) )  - \prob( B_i \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) > D_i) \prob(B_i = b) - \sum_{t=1}^{\min\sched_i} \prob(B_i = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(t_i)+1}^T (1 - S_\theta(\tau_{\sched_i}(b + 1))) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_\theta(\tau_{\sched_i}(b + 1))) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_\theta(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(t_i)+1}^T S_\theta(\tau_{\sched_i}(b + 1)).
\end{align}

@@

\subsection{Results}

>>>>

\subsection{Approach} \label{perf-test:sec:approach}

@@

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$.
  Otherwise, define $l_i^{(b)}$ and $r_i^{(b)}$ as the first and last days $b_i$ could have occurred.
  $l_i^{(b)}$ is the day after the last negative test prior to the first positive test in the episode.
  $r_i^{(b)}$ is the day of the first positive test.
  Similarly, define $l_i^{(e)}$ and $r_i^{(e)}$ as the first and last days $e_i$ could have occurred.
  If $i \in \set{D}$ then $n_{ia} =1$ admissible episodes occurred in individuals with test schedule $\sched_{i}$, otherwise $n_{ia} = 0$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is admissible with probability $p_{ia} = \prob((b, e) \in \alpha_i \mid \theta)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i^C$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these observed in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, it is undetected with probability $p_{iu} = \prob((b, e) \in \Omega^C_i \mid \theta)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  This is all remaining episodes not in the previous sets.
  $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, it is inadmissible with probability $p_{i\bar{a}} = \prob((b, e) \in \beta_i \mid \theta)$.
\end{itemize}

@@

Next, I derive $p_{ia}$.
By definition (see~\cpageref{perf-test:def:admissible}), we have:
\todo{Make sure "detectable" is defined: It's really a latent quantity, which could possibly be defined in terms of true viral load. I think "positive probability of testing positive" under the assumption of no false positives is what I will go with, or by reference to a multi-state model type view of infection.}
\begin{align}
p_{ia} &= \prob((b, e) \in \alpha_i \mid \theta) \\
\alpha_i &= \{ (b, e) : l_i^{(b)} \leq b \leq r_i^{(b)} \wedge l_i^{(e)} \leq e \leq r_i^{(e)}\}.
\intertext{Suppressing the conditioning on $\theta$, this gives:}
p_{ia}
&= \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)}, l_i^{(e)} \leq E_i \leq r_i^{(e)} \right) \\
&= \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \prob \left( l_i^{(b)} \leq B_i \leq r_i^{(b)} \right) \\
&=\sum_{b = l_i^{(l)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} \leq E_i \leq r_i^{(e)} \mid B_i = b \right) \prob \left(B_i = b \right) \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \prob \left( l_i^{(e)} - b + 1 \leq D_i \leq r_i^{(e)} - b + 1 \right) \prob \left(B_i = b \right) &\text{by def of $D_i$} \\
&=\sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right) \prob \left(B_i = b \right) &\text{by def of $S_\theta$} \\
&\propto \sum_{b = l_i^{(b)}}^{r_i^{(b)}} \left( S_\theta(l_i^{(e)} - b + 1) - S_\theta(r_i^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.

Next, I derive $p_{iu} = \prob((B_i, E_i) \in \Omega^C_i \mid \theta)$, the probability that episode $i$ was undetected.
An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
Then $\Omega^C_i$ can be written as:
\begin{align}
\Omega_i^C = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
\end{align}
Only $1 - p_u = 1 - \frac{1}{N} \sum_i^{\Ncis} p_{iu} = \frac{1}{N} \sum_i^{\Ncis} (1 - p_{iu})$ is required (see \cref{perf-test:eq:full-posterior-simplified}).
Therefore, I derive $1 - p_{iu}$, again suppressing the conditioning on $\theta$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_i, E_i) \in \Omega^C_i) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) + B_i > E_i \vee B_i \leq \min(\sched_i) ) \\
&= 1 - \prob( \tau_{\sched_i}(B_i) + 1 > D_i , B_i > \min(\sched_i) )  - \prob( B_i \leq \min(\sched_i) ) \\
&= 1 - \sum_{b={\min\sched_i} + 1}^T \prob( \tau_{\sched_i}(b) + 1> D_i) \prob(B_i = b) - \sum_{t=1}^{\min\sched_i} \prob(B_i = b)\\
&= 1 - \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \frac{1}{T} \min\sched_i \\
&= \frac{1}{T} \left(T - \sum_{b=\min\sched_i+1}^T (1 - S_\theta(\tau_{\sched_i}(b) + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \left(T - (T - \min\sched_i) + \sum_{b=\min\sched_i+1}^T S_\theta(\tau_{\sched_i}(b + 1)) - \min\sched_i \right) \\
&= \frac{1}{T} \sum_{b=\min(\sched_i)+1}^T S_\theta(\tau_{\sched_i}(b) + 1). \label{perf-test:eq:piu}
\end{align}

@@

\subsection{Results} \label{perf-test:sec:results}
